### 2018-04-18 PRESENTATIONS/presentation.Rmd
paa presentation

---
title: "Climate change projected to reduce European life expectancy"
author: | 
 | Mathew E. Hauer^1^ & Alexis R. Santos^2^
 | ^1^University of Georgia | ^2^Pennsylvania State University
date: "April 26, 2018"
output:   beamer_presentation
classoption: "aspectratio=169"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
library(tidyverse)
library(ggrepel)
library(grid)
library(gridExtra)
library(HMDHFDplus)
library(scales)
library(dplyr)
library(openxlsx) 
library(tmap)
library(tmaptools)
library(tigris)
library(censusapi)
library(tidycensus)
library(cowplot)
library(magick)
library(pdftools)
library(readxl)
```

```{r setup3, include=FALSE, cache =TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(png)
rm(list=ls())
library(tidyverse)
library(HMDHFDplus)
library(tmap)
library(tmaptools)
library(RColorBrewer)
library(data.table)
library(getPass)

###   Inputting user info for HMD

myusername <- getPass(msg = "HMD username: ", noblank = FALSE, forcemask = FALSE)
mypassword <- getPass(msg = "HMD password: ", noblank = FALSE, forcemask = FALSE)

###   Importing the data from Forizeri et al.
lancetdata <- tribble(
  ~CNTRY,~MID,~LOW,~HIGH,
  "AUT",467,121,797,
  "BEL",2841,91,5007,
  "BGR",44,36,63,
  "CHE",1981,497,3247,
  "CYP",143,62,279,
  "CZE",375,48,628,
  "DEUTNP",7319,620,12132,
  "DNK",24,4,56,
  "EST",12,0,22,
  "ESP",41326,33495,53802,
  "FIN",23,3,44,
  "FRATNP",38728,14235,72947,
  "GRC",2838,1520,5606,
  "HRV",1145,624,1811,
  "HUN",559,191,985,
  "IRL",1566,476,4195,
  "ISL",3,1,5,
  "ITA",41965,23694,57401,
  "LTU",15,1,25,
  "LUX",392,12,820,
  "LVA",17,1,30,
  "MLT",0,0,0,
  "NLD",1984,126,4114,
  "NOR",15,4,34,
  "POL",313,52,561,
  "PRT",4573,3808,6284,
  "ROU",177,97,301,
  "SWE",52,10,113,
  "SVN",561,251,847,
  "SVK",63,17,118,
  "GBR_NP",2515,388,7484
)

# ec_agestandard <- tribble(
#   ~Age, ~StnrdPop,
#   0, 1600,
#   1, 6400,
#   5, 7000,
#   10,7000,
#   15,7000,
#   20,7000,
#   25,7000,
#   30,7000,
#   35,7000,
#   40,7000,
#   45,7000,
#   50,7000,
#   55,6000,
#   60,5000,
#   65,4000,
#   70,3000,
#   75,2000,
#   80,2000
# )

ec_agestandard <- tribble(
  ~Age, ~StnrdPop,
  0, 1118.427,
  1, 4338.143,
  5, 5207.188,
  10, 5378.67,
  15, 6095.53,
  20, 6646.578,
  25, 7054.462,
  30, 7211.366,
  35, 7249.137,
  40, 7288.966,
  45, 7207.381,
  50, 6904.728,
  55, 6400.144,
  60, 5798.191,
  65, 4660.589,
  70, 4031.35,
  75, 3292.724,
  80,4116.422
)

europe_standarddr <- tribble(
  ~COUNTRY, 	~Circulatorydisease, 	~Cancer, 	~Heartdisease, 	~LungCancer, 	~Respiratorydiseases, 	~Diseasesofthenervoussystem, 	~Colorectalcancer, 	~Suicide, 	~TransportAccidents, 
  'EU-28', 	373.6, 	261.5, 	126.3, 	54.4, 	78.3, 	38.6, 	30.5, 	11.3, 	5.8, 
  'Luxembourg', 	296.9, 	260.7, 	80.3, 	59.6, 	63.8, 	38, 	25.5, 	13.4, 	6, 
  'Spain', 	245, 	232.7, 	68.2, 	47.8, 	91.7, 	48.5, 	33.6, 	8.2, 	4.3, 
  'Italy', 	310.1, 	246.6, 	98.3, 	49.4, 	58.3, 	34.3, 	27, 	6.3, 	5.6, 
  'France', 	202.9, 	245.4, 	49.3, 	50.1, 	52, 	50.2, 	26.1, 	14.1, 	5.1, 
  'Portugal', 	305.8, 	242.1, 	69.6, 	36.4, 	116.7, 	32.8, 	35, 	11.3, 	7.8, 
  'Croatia', 	678.6, 	336.4, 	306.5, 	65.2, 	59.7, 	21.3, 	51, 	16.8, 	8.9, 
  'Ireland', 	309.9, 	288.3, 	147.5, 	61.5, 	125.9, 	48.7, 	32.4, 	11, 	4, 
  'Slovenia', 	451.3, 	299.9, 	102.8, 	58.6, 	66.3, 	21.1, 	38.4, 	18.9, 	6.7, 
  'Switzerland', 	280, 	219.6, 	97.8, 	42.1, 	51.3, 	44.5, 	22.8, 	12.8, 	3.6, 
  'Belgium', 	281.9, 	252.6, 	72.4, 	61.6, 	95.7, 	46.5, 	26.1, 	17.3, 	6.7, 
  'Greece', 	381.4, 	249.3, 	103, 	61.9, 	108.1, 	20.9, 	23.3, 	5, 	8.6, 
  'Netherlands', 	271.7, 	282.2, 	62.4, 	66.7, 	74.1, 	48.3, 	32.9, 	11.1, 	4.1, 
  'Germany', 	403.5, 	253.2, 	142.8, 	51, 	68, 	29.6, 	29, 	11.9, 	4.6, 
  'Great Britain', 	264.9, 	278.4, 	118.4, 	61.4, 	130.9, 	47.6, 	27.7, 	7.1, 	2.8, 
  'Austria', 	418.1, 	249.3, 	179.1, 	47.5, 	46.6, 	32.6, 	26.4, 	15.3, 	5.8, 
  'Bulgaria', 	1131, 	242.4, 	195.4, 	45.5, 	58.1, 	15.3, 	34.9, 	9.9, 	9, 
  'Czech Republic', 	615.2, 	284.6, 	333.1, 	53.1, 	73.4, 	30.8, 	37.9, 	14.4, 	78.8, 
  'Denmark', 	256.6, 	300.6, 	81, 	71.7, 	115.7, 	42.9, 	35.2, 	11.9, 	4, 
  'Estonia', 	699.6, 	299.4, 	295.5, 	55.3, 	43.8, 	21.8, 	36, 	18.3, 	7.5, 
  'Finland', 	378.8, 	218.6, 	199.2, 	39, 	34.4, 	155, 	22.6, 	14.6, 	5.7, 
  'Hungary', 	761.5, 	348.1, 	390.6, 	89.8, 	78.6, 	19.9, 	55, 	19.4, 	8.1, 
  'Latvia', 	882.7, 	299.3, 	442.7, 	46.9, 	35.9, 	15.6, 	34.2, 	19, 	12.4, 
  'Lithuania', 	848.8, 	276.2, 	564.4, 	46.1, 	42.1, 	20.8, 	32.1, 	31.5, 	10.7, 
  'Norway', 	272.6, 	252.5, 	95.7, 	50.5, 	88.4, 	45.4, 	36.4, 	7.3, 	4, 
  'Poland', 	591.4, 	292.3, 	129.1, 	69.2, 	69.1, 	16.5, 	36, 	15.5, 	10.3, 
  'Slovakia', 	654.6, 	324.1, 	388.8, 	50, 	74.9, 	29.5, 	49.2, 	10.8, 	8.5, 
  'Sweden', 	338.3, 	234.8, 	131.2, 	38.7, 	58.1, 	42.6, 	29.2, 	12.1, 	3.4, 
  'Cyprus', 	351.8, 	201, 	108.7, 	37.2, 	86.2, 	26.8, 	16.7, 	4.5, 	6.5, 
  'Liechtenstein', 	296.4, 	203, 	73.7, 	31.3, 	89.8, 	67.6, 	6.8, 	10.2, 	10.3, 
  'Malta', 	372.4, 	233.5, 	202.8, 	43.2, 	96.6, 	21, 	28.3, 	8.3, 	2.5, 
  'Romania', 	951.3, 	273.2, 	320.3, 	54.2, 	78.4, 	21, 	32.4, 	11.4, 	12.3, 
  'Serbia', 	931.6, 	298.3, 	159.5, 	69.4, 	79.7, 	27.3, 	37.2, 	15.9, 	7.6

)

countrycodes <- read.csv("https://raw.githubusercontent.com/lukes/ISO-3166-Countries-with-Regional-Codes/master/all/all.csv") %>%
  rename(CNTRY = alpha.3,
         COUNTRY = name) %>%
  mutate(ISO3 = CNTRY) %>%
  select(COUNTRY, CNTRY, ISO3)
#  read.csv("data/LANcET_CCMORTDATA.csv",colClasses=c("CNTRY"="character"))

###   Getting a country list from the HMD
Countries <- getHMDcountries()

###   Downloading the DEATHS data from the HMD in 5-year age groups by single-year.
deaths <- rbindlist(lapply(Countries, function(CNTRY){
  Dat <- readHMDweb(CNTRY = CNTRY, item = "Deaths_5x1", fixup=TRUE, username = myusername, password = mypassword)
Dat$CNTRY <- CNTRY
Dat}))

###   Downloading the POPULATION data from the HMD.
pops <- rbindlist(lapply(Countries, function(CNTRY){
  Dat <- readHMDweb(CNTRY = CNTRY, item = "Population", fixup=TRUE, username = myusername, password = mypassword)
  Dat$CNTRY <- CNTRY
  Dat}))

###   Downloading the LIFE TABLE data from the HMD in 5-year age groups by single-year.
lt <- rbindlist(lapply(Countries, function(CNTRY){
  Dat <- readHMDweb(CNTRY = CNTRY, item = "bltper_5x1", fixup=TRUE, username = myusername, password = mypassword)
  Dat$CNTRY <- CNTRY
  Dat}))

###   Topcoding the age groups to 80+ of the LIFE TABLE data, and summing the ax value.
lt2 <- lt%>%
  mutate(Age = ifelse(Age >= 80, 80, Age)) %>%
  group_by(CNTRY, Year, Age)%>%
  summarise(ax = sum(ax))

###   Topcoding the age groups of the DEATHS data and summing.
deaths2 <- deaths %>%
  mutate(Age = ifelse(Age >=80, 80, Age)) %>%
  group_by(CNTRY, Year, Age) %>%
  summarise(deaths = sum(Total) )

###   Recoding the Population data to single year of age
pops2 <- pops %>%
  mutate(Age = ifelse(Age >= 80, 80, 
               ifelse(Age >= 75, 75,
               ifelse(Age >= 70, 70,
               ifelse(Age >= 65, 65,
               ifelse(Age >= 60, 60,
               ifelse(Age >= 55, 55,
               ifelse(Age >= 50, 50,
               ifelse(Age >= 45, 45,
               ifelse(Age >= 40, 40,
               ifelse(Age >= 35, 35,
               ifelse(Age >= 30, 30,
               ifelse(Age >= 25, 25,
               ifelse(Age >= 20, 20,
               ifelse(Age >= 15, 15,
               ifelse(Age >= 10, 10,
               ifelse(Age >= 5,  5,
               ifelse(Age >= 1,  1, Age)))))))))))))))))) %>%
  group_by(CNTRY, Year, Age) %>%
  summarise(Pop= sum(Total1))

###   Joining the POPULATION, DEATHS, and LIFE TABLE data and subseting to the maximum year in the data.
a <- left_join(pops2, deaths2)
a <- left_join(a, lt2)
a <- group_by(a, CNTRY, Age) %>%
  filter(Year == max(Year))

###   Importing the GBD data.
###   Data is 5-year age groups of mortality rates for Environmental heat and cold exposure for 2006-2015. 
###   Analysis uses the mean of the last 10 years of data for this death distribution.
GBD_data <- read.csv("../data/GBD_data.csv") %>%
  #mutate(perdist = log(val)) %>%
  group_by(countrycode, Age) %>%
  summarise(meanval = mean(val),
            upperval = mean(upper),
            lowerval = mean(lower)) %>%
  group_by(countrycode) %>%
  mutate(countmeanval = sum(meanval),
         countupperval = sum(upperval),
         countlowerval = sum(lowerval),
         perdist = (meanval / countmeanval)) %>%
  rename(CNTRY = countrycode)

ggplot(data=GBD_data, aes(x=Age, y=perdist)) +
  geom_line(aes(group =CNTRY, col=CNTRY), alpha=0.7) +
  geom_smooth(span=0.3, se=F, col="black") +
  theme_bw() +
  scale_y_log10()+
  labs(y="log(Mortality Percentage Distribution)")

a <-  left_join(a, lancetdata)  %>%
  select(Age, Pop, CNTRY, deaths, ax, MID, LOW, HIGH) %>%
   filter(!is.na(MID))
a2 <- left_join(a, GBD_data) %>%
  mutate(width = ifelse(Age == 0, 1,
                 ifelse(Age == 1, 4,5)),
         mx_base = deaths/Pop,
         mx_low = (deaths + LOW * perdist)/Pop,
         mx_mid = (deaths + MID * perdist)/Pop,
         mx_high = (deaths + HIGH * perdist)/Pop,
         qx_base = ifelse(Age == 80, 1, mx_base/(1+((width-ax)*mx_base))),
         qx_low = ifelse(Age == 80, 1, mx_low/(1+((width-ax)*mx_low))),
         qx_mid = ifelse(Age == 80, 1, mx_mid/(1+((width-ax)*mx_mid))),
         qx_high = ifelse(Age == 80, 1, mx_high/(1+((width-ax)*mx_high)))) %>%
  group_by(CNTRY) %>%
  mutate(lx_base = ifelse(is.na(lag(cumprod(1-qx_base),1)*100000), 100000, lag(cumprod(1-qx_base),1)*100000),
         dx_base = qx_base*lx_base,
         Lx_base = ifelse(Age == 80, (lx_base/mx_base), ax * lx_base + (width-ax) * lead(lx_base,1)),
         Tx_base = rev(cumsum(rev(Lx_base))),
         ex_base = Tx_base/lx_base,
         lx_low = ifelse(is.na(lag(cumprod(1-qx_low),1)*100000), 100000, lag(cumprod(1-qx_low),1)*100000),
         dx_low = qx_low*lx_low,
         Lx_low = ifelse(Age == 80, (lx_low/mx_low), ax * lx_low + (width-ax) * lead(lx_low,1)),
         Tx_low = rev(cumsum(rev(Lx_low))),
         ex_low = Tx_low/lx_low,
         lx_mid = ifelse(is.na(lag(cumprod(1-qx_mid),1)*100000), 100000, lag(cumprod(1-qx_mid),1)*100000),
         dx_mid = qx_mid*lx_mid,
         Lx_mid = ifelse(Age == 80, (lx_mid/mx_mid), ax * lx_mid + (width-ax) * lead(lx_mid,1)),
         Tx_mid = rev(cumsum(rev(Lx_mid))),
         ex_mid = Tx_mid/lx_mid,
         lx_high = ifelse(is.na(lag(cumprod(1-qx_high),1)*100000), 100000, lag(cumprod(1-qx_high),1)*100000),
         dx_high = qx_high*lx_high,
         Lx_high = ifelse(Age == 80, (lx_high/mx_high), ax * lx_high + (width-ax) * lead(lx_high,1)),
         Tx_high = rev(cumsum(rev(Lx_high))),
         ex_high = Tx_high/lx_high,
         DIF_LOW = ex_low - ex_base,
         DIF_MID = ex_mid - ex_base,
         DIF_HIGH = ex_high - ex_base)

# write.table(a2, "data/r_output2.txt", sep="\t")

lt_lancet <- a2 %>%
  filter(Age == 0) %>%
  mutate(Base_e0 = ex_base,
         LOW_e0 = ex_low,
         MID_e0 = ex_mid,
         HIGH_e0 = ex_high) %>%
  #group_by(CNTRY) %>%
   dplyr::select(CNTRY, Base_e0, LOW_e0, MID_e0, HIGH_e0, DIF_LOW, DIF_MID, DIF_HIGH) %>%
  ungroup() %>%
 arrange(DIF_MID) %>%
  mutate(RANK = row_number(),
         dif_mid = round(DIF_MID, 2)) %>%
  left_join(., countrycodes) %>%
  mutate(COUNTRY = case_when(
    CNTRY == "FRATNP" ~ "France",
    CNTRY == "DEUTNP" ~ "Germany",
    CNTRY == "GBR_NP" ~ "Great Britain",
    TRUE ~ as.character(COUNTRY)),
    ISO3 = case_when(
      CNTRY == "FRATNP" ~ "FRA",
      CNTRY == "DEUTNP" ~ "DEU",
      CNTRY == "GBR_NP" ~ "GBR",
      TRUE ~ as.character(ISO3)),
    Test = ISO3)

```

## Climate Change and Human Health

![](./presentation_files/figure-beamer/climate_change_health_impacts600w.jpg)\

morbidity $\neq$ mortality

## Climate Change and Mortality

Typically described as excess or "extra" mortality. Most studies quantify exposure rather than mortality.

Some studies do publish projected mortality:

- Forzieri et al (Europe), six-climate hazards: 150k deaths by 2080s
- Silva et al (Global), ozone-related mortality: 375k by 2100

These estimates are useful but sterile -- *one death is a tragedy, a million deaths a statistic*.

## Climate Change and Life Expectancy

Life expectancy at birth ($e_0$) is the preferred mortality metric.

To our knowledge, *no studies quantify the possible impact of climate change on human longevity*.

Without properly situating the potential loss of life within metrics like life expectancy, we risk miscommunicating the impact of climate change on human mortality.

We ask and answer the following questions:

- What is the cost of inaction on climate change on human longevity?
- How does climate change related mortality compare to other mortality vectors?

## Data 1

We use data from three primary sources:

1. Forzieri et al projected excess mortality due to six climate vectors (heatwaves, coldwaves, wildfires, droughts, river and coastal floods, and windstorms) under conventional greenhouse gas emission scenario (SRES A1B). (150k deaths per year by 2080s)

![](./presentation_files/figure-beamer/forzieri.PNG){ width=45% } via: Forzieri et al (2017)

## Data 2

2. To convert Forzieri et al's excess mortality to life expectancy, we use the Human Mortality Database (HMD) for underlying life tables.

- HMD for the correspoding 28 European countries in 5x1 (by age and year) life tables for the most recent life table.

    - Abridge the life table to $[0,1,5,...,80+]$

- $_nP_x$, population aged $x$ to $x+n$

- $_na_x$, average length of survival between ages $x$ and $x+n$

- $_nd_x$, number of deaths between ages $x$ and $x+n$

## Data 3

3. The Global Burden of Disease (GBD) provides cause-specific mortality by age/sex/geography/year for 249 causes of death.

- We gathered ata on age-specific deaths and mortality rates for environmental heat and cold exposure (cause C.2.9) for 2006-2015 for each country in the study.

```{r suppfig, echo= FALSE, message = FALSE, warning = FALSE, fig.width = 4, fig.height=2, fig.align = "center"}
GBD_data <- read.csv("../data/GBD_data.csv") %>%
  #mutate(perdist = log(val)) %>%
  group_by(countrycode, Age) %>%
  summarise(meanval = mean(val),
            upperval = mean(upper),
            lowerval = mean(lower)) %>%
  group_by(countrycode) %>%
  mutate(countmeanval = sum(meanval),
         countupperval = sum(upperval),
         countlowerval = sum(lowerval),
         perdist = (meanval / countmeanval)) %>%
  rename(CNTRY = countrycode)

ggplot(data=GBD_data, aes(x=Age, y=perdist)) +
  geom_line(aes(color=CNTRY), alpha=1) +
  geom_smooth(span=0.3, se=F, col="black", lwd =2) +
  theme_bw() +
  theme(legend.position = "none",
        legend.title=element_blank(),
        legend.text=element_text(size=2),
        legend.key.size = unit(1,"line")) +
  # guides(col = guide_legend(ncol = 4)) +
  scale_y_log10()+
  labs(y="log(Mortality)")
       #title = "Mortality distribution of Heat-related mortality in 28 European countries")

```

## Methods

- From the GBD data, we derived variable $t_{x,i}$ as the proportion of deaths, $D$, from each age group $x$ in each country $i$ ($_nt_{x,i}=D_{x,i}^{GBD}/\sum_{\alpha=0}^{80}{_nD_i^{GBD}}$).

- We then derived additional $m_x$ rates for each age group $x$ in each country $i$ for each scenario $s$ ($BASE$, $LOW$, $MID$, $HIGH$) from Forzieri et al.

\begin{equation}
_n\hat{m}_{x,i,s} = (_nD_{x,i}^{HMD} + \mathbf{(\hat{D_i} \,\cdot\, _nt_{x,i})} ) \,/\, _nP_{x,i}^{HMD}
\end{equation}

- Calculate the remaining life table using standard life table techniques and simply subtract $e_{x,i,s}$ from $e_{x,i,BASE}$ to arrive a $\Delta e_x$.

## Results

```{r results1, echo= FALSE, message = FALSE, warning = FALSE, fig.align = "center"}
ggplot(lt_lancet, aes(x=RANK, y=DIF_MID, ymin = DIF_LOW, ymax = DIF_HIGH, label=dif_mid)) +
  geom_pointrange(stat="identity") +
  geom_point(stat="identity", fill="black", size=15) +
   geom_hline(yintercept=0) +
  geom_text(color="white", size=4.5) +
  #scale_x_discrete(limits = rev(levels(the_factor)))
  scale_x_reverse(breaks = lt_lancet$RANK, labels=lt_lancet$COUNTRY) +
  coord_flip() +
  theme_bw() +
  #ylim(-2, 2) +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size=20)) +
    labs(y=expression(Delta* e[0]),
        x= "Country")

#, fig.width = 10, fig.height=6,

```

## Results
```{r results2, echo= FALSE, message = FALSE, warning = FALSE, fig.align = "center"}
data(Europe)
Europe2 <- append_data(Europe, lt_lancet, key.shp = "iso_a3", key.data = "ISO3", ignore.na = TRUE)

bbeur <- bb(Europe2, xlim = c(0,0.75), ylim=c(0,1), relative = TRUE)

  tm_shape(Europe2) +
  tm_fill("DIF_MID", title = expression(Delta* e[0 ~italic(MID)]), palette = "YlOrBr", style = "jenks") +
  tm_borders(alpha = 0.5) +  
  tm_text("Test" , size="AREA", root=5) +
  tm_format_Europe()

```

## Results

![](./presentation_files/figure-beamer/table2.PNG){ width=100% }\

## Conclusion

- Two European countries could see reductions of more than one-year ($MID$) but five countries could see more than one-year reduction under higher scenarios ($high$).

- Thus, **the cost of inaction on climate change could approach *one year of life* in some countries**.

- Life expectancy connects mortality estimates into more intuitively understandable metrics, translating global estimates of mortality into individual outcomes.

- Without adaptation measures, our results suggest climate change could emerge as a significant new mortality vector for some European countries by the end of the century.

## Thank you!

**Mathew E Hauer**
Applied Demographer 
Carl Vinson Institute of Government \
hauer@uga.edu

**Alexis R Santos **
Applied Demographer
Pennsylvania State University